---
title: "10_17_Polynomials"
output: 
  html_document:
    keep_md: true
---

# Polynomials 
## Chapter 5.1 in Sheather 

Common R Command: 
  - poly 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
library(rms)
```

```{r}
attach(cars)
plot(size,mpg)
size2<-size^2 
m<-lm(mpg~size+size2)
plot(size,mpg)
points(size, m$fit, pch=16, col="orange")
## The importance of centering
sizec2<-(size-mean(size))^2
cor(cbind(size,size2, sizec2))
bad<- 900:910
bad
bad^2
plot(bad, bad^2)
plot(bad - 900, (bad-900)^2)

plot(size,mpg)
points(size,m$fit, pch=16, col="orange")
summary(m)
m<-lm(mpg~size+sizec2)
points(size, m$fit, pch=16, col="blue")

sizeo2<- lm(size2~size)$residual
par(mfrow=c(1,3))
plot(size, size2)
plot(size, sizec2)
plot(size, sizeo2)
m<- lm(mpg~size+sizeo2)
summary(m)

plot(size,mpg)
points(size, m$fit, pch=16, col="green")

```

## Fitting a cubic polynomial to the data 
$ y = \beta_{0} + \beta_{1}x + \beta_{2}x^2 + \beta_{3}x^3$ 
Actually, because of centering the values looks like:
$ y = \beta_{0} + \beta_{1}x + \beta_{2}(x- \bar{x})^2 + \beta_{3}(x-\bar{x})^3$


```{r}
sizec3<- (size-mean(size))^3
m<-lm(mpg~size+sizec2+sizec3)
summary(m)

plot(size, mpg)
points(size, m$fit, pch=16, col="blue")
```
## The poly function 
Rescales it so it's not too big and takes the orthogonal polynomial. Very numerically stable. 

```{r}
poly(size,2)
x<-1:1000
matplot(x, poly(x,2), type="l")
matplot(x, poly(x,3), type="l")
matplot(x, poly(x,4), type="l")

m<-lm(mpg~poly(size,2))
summary(m)

m<-lm(mpg~poly(size, 3))
summary(m)

m<-lm(mpg~poly(size, 4))
summary(m)

plot(size,mpg)
points(size, lm(mpg~poly(size,4))$fitted, pch=16, col="red")
m<-lm(mpg~size+sizec2+sizec3)
summary(m)

anova(m)
```

Looking at the ANOVA table for the polynomial models helps one see how much you gain by adding the additional terms. In the ANOVA(m) above, the cubic term is adding about 1/20 of what the regression added, so it's making a subtle adjustment. 

###Another example: 

Notice the linear spline (a knot that ties together to simple functions (i.e. lines)) that occurs at Age = 1.5. 

```{r}
head(vocabulary)
attach(vocabulary)
plot(Age,Vocab)
kink(Age, 1.5)
m<-lm(Vocab~Age+kink(Age, 1.5))
summary(m)

plot(Age,Vocab)
points(Age,m$fit, pch=16)


#plot(size, mpg)
#points(size, lm(mpg~lsp(size, 200))$fitted, pch=16)
#points(size, lm(mpg~lsp(size, c(100, 250)))$fitted, pch=16)
```

### Linear Spline Function: 
$ (x-k_{j})_{+}$  - positive part of $x-k_{j}$ where k is the knot. 

In general $ y = \beta_{0} + \beta_{1}x + \beta_{2}(x- k_{1})^2 + \beta_{3}(x-k_{2})^3 +...+ \beta_{j+1}(x-k_j)_+ +\epsilon$

A Cubic Spline:  $ y = \beta_{0} + \beta_{1}x + \beta_{2}x^2 + \beta_{3}x^3 + \beta_4(x-k_4)^3...+ \beta_{n+3}(x-k_n)^3_+ +\epsilon$

A restricted cubic spline or a natural spline: at the last knot, it smoothes the function into a line. Don't add a cubic at the last step, add a linear element. 

```{r}
x<-rcspline.eval(size, nk=3, inclx=TRUE)
matplot(size,x, type="l")


x<-rcspline.eval(size, nk=4, inclx=TRUE)
matplot(size,x, type="l")


```



