---
title: "Notes, Tuesday, August 27: Review of Basic Statistics"
output: html_notebook
---

# Tuesday, August 27 
## Review of Basic Statistics 

```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
```
```{r}
install.packages("ltm")
library(ltm)
install.packages("car")
library(car) 
```

x(1), x(2), ... x(n)

mean = [x(1) + x(2)+...+x(n)]/n = $\sum_{x = 1}^{n} f(x(i))$/n


Basic commands
mean(variable): 
sort(variable): lowest-> hightest
median(variable)

mean & median are measures of location: they will move x distance when +/- x. 
mean is more likely to be skewed by outliers, most of statistics built on the mean however which can produce issues.

summary(variable): min, max, median, mean, quartiles (also a measure of location) 

low<- variable<=median
variable[low] : pull out values lower than the median
variable[!low]: pull out values not (!) than the median 

boxplot(variable): plots a boxplot (information from summary)

sample variance: $s^2$ = $\sum_{x = 1}^{i} f(x(i) -mean)$/n-1 

Introduction to probability. Flip a coin, 1000 times

#