---
title: "8/29 Basic Probability"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
```

# Thursday, August 29 
## Basic Probability 


**Sample Space:** Set of everything $\Omega$ that might happen 

**Random Vairable:** Rule or function that assignes a number to each element of $\Omega

**Independence:** Independence of 2 random variables X and Y: Pr(X=x and Y=y)=Pr(X=x)Pr(Y=y)

**Variance** Variance or how far is x from the expectation of x: variance= $E[(X-E(X))^2]$

**Standard Deviation:** $\sqrt(variance)$

**Density Function:** f(x) of density function has two properties $f(x) \ge 0$ and total area under f(x)=1

**Central Limit Theorem:** mean of n nice independent random vairable is nearly normal as n approaches $\infty$


```{r}
red<-sample(c("head", "tail"), 10000, replace=TRUE)
blue <- sample(c("head","tail"), 10000, replace=TRUE)
red[1:10]
blue[1:10]
cbind(red,blue)[1:10,]
table(red,blue)
y<-1*(blue=="head")
table(y, blue)
x<-1*(red=="tail")
table(x,y)
addmargins(table(x,y))
toss10<- rbinom(10000,10,.5)
toss10[1:30]
table(toss10)
barplot(table(toss10))
```



```{r}
data(Abortion)
head(Abortion)
help(Abortion)
apply(Abortion,2,mean)
table(Abortion[,1], Abortion[,3])

```

Picking a random number 

```{r}
sample(0:9,1)
```

**Density Functions:**
*Normal density* is not the normal density- it's not particularly normal. Sometimes called the Gaussian distribution. 

X~N(0,1) -> expectation 0, variance 1
$y= \sigma X +\mu$ ~$N(\mu, \sigma^2)$ -> expectation $\mu$, variance $\sigma^2$

Normal distribution has 2 roles: as a model for Data and in the Central limit theorem 

mean of n independent Y(i) ~ $N(\mu, \sigma^2)$ is $N(\mu, \sigma^2/n)$


```{r}
plotnormal(0,1)
abline(v=0, col="blue")
pnorm(1)
rnorm(100)
mean(rnorm(100))
mean(rnorm(1000))
```


*Cauchy Distribution* Y is Cauchy then the mean of n indepentdent Caushy's is Cauchy (???)

```{r}
plotcauchy(0,1)
rcauchy(100)
mean(rcauchy(100))
mean(rcauchy(10000))

```