---
title: "9_10_Simple_Regression"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
```

# Tuesday, September 10
## Simple Regression (one X)
### Properties of Least Squares estimates 
 - unbiased
 - minimum variance unbiased
 - not robust 
 



```{r}
head(fuel)
attach(fuel)
arrowplot(Tax,Fuel)
m<- lm(Fuel~Tax)
m$residual
summary(m)

dogdata
attach(dogdata)
boxplot(Change)
qqnorm(Change)
qqline(Change)
shapiro.test(Change)

```

```{r}
par(mfrow=c(2,2))
fuelsim()
fuelsim()
fuelsim()
fuelsim()

fs <- fuelsimbeta(1000)
boxplot(fs) 
abline(h=-50)

mean(fs)

sd(fs)

set.seed(88)
gm<-fuelsimbeta(1000)
sd(gm)

summary(m)

```


Fitted line- least square fit 
True Line : $Y_{i}=\alpha +\beta X_{i} + E_{i}$
 
 
 Under the **normal model** least squares estimates are unbiased. 
 -$E(\widehat{\beta})= \beta$
 -$E(\widehat{\alpha})= \alpha$
 -$E(\widehat{\sigma}^2)= \sigma^2$
 
 Want the variability of $\widehat{\beta}$ to be small. Thus compute the variation or SD of $\widehat{\beta}$ 
 
 $var(\widehat{\beta})=E[(\widehat{\beta}-E(\beta)^2)]= E((\widehat{\beta}-\beta)^2)$
 or in words, Variance of estimate = expected value of the squared difference of estimate - parameter
 
 ### **Gauss-Markov Theorem**
 Among all unbiased estimates of $\beta$, the least squares estimate has the smallest variance. 
 
 $var(\widehat{\beta})=\frac{\sigma^2}{\sum_{i=1}^n(x_{i}-\bar{x})^2}$
 
This forumla tells you it's hard to estimate $\beta$ when there's a lot of variation. But it's good to estimate $\beta$ over a wide range. However you want most of the data points to account for the variabilty in x.You don't want the variability to be coming from just a few outlining data points in x. 
 
If the normal model is true: 
- $\hat\beta$ is normal 
- with mean $\beta$ 
- and variance, $\frac{\sigma^2}{\sum_{i=1}^n(x_{i}-\bar x)^2}$

95% Confidence Interval for $\beta$
$\hat\beta \pm$ allow for error
$\hat\beta \pm t_{.025} * \hat{se}(\hat\beta)$

```{r}
qt(.975,46)
confint(m)
-53.11+2.012*15.48

```

### Hypothesis Testing 
$H_{0}: \beta = \beta_{0}$
most commonly, $\beta_{0}=0$
$t=\frac{\hat\beta -\beta_{0}}{\hat{se}(\hat\beta)}$

### Making Predictions: New X', what should I predict? 
 - Where is $\alpha +\beta X'$
 - Where is a new observation" $\alpha+ \beta X' + \epsilon$
```{r}
plot(Tax,Fuel)

```
 