---
title: "9_17_Correlation"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
```



# Tuesday, September 17 
## Correlation 

```{r}
attach(fuel)
cor(Fuel, Tax)
```
** Pearson's Correlation** 
- Oldest form of correlation and most tied to linear regression, not necessarily the most intuitive. If type of correlation is unspecified, it's pearson's correlation. 

Correlation is between -1 and 1. 


```{r}
Fuels<- (Fuel-mean(Fuel))/sd(Fuel)
Taxs <- (Tax-mean(Tax))/sd(Tax)
mean(Fuels)
sd(Fuels)

plot(Tax, Taxs)
summary (Tax)
summary (Taxs)

par(mfrow=c(1,2))
plot(Tax, Fuel)
plot(Taxs, Fuels)
lm(Fuel~Tax)
lm(Fuels~Taxs)
abline(lm(Fuels~Taxs), lwd=2, col="red")

head(corEG)
attach(corEG)
cor.test(Fuel, Tax)
```
How big are the values in terms of the mean and standard deviation. 

## What correlation doesn't explain 
```{r}
cor(x,y3)
plot(x,y3)
abline(lm(y3~x), col="blue", lwd=2)
```


Pearson's Correlation only explains when variables are related by a line. Thus it gives us 0 for the plot of x, y3 above. Even though we know they are perfectly related. 

```{r}
cor(x, y4)
plot(x, y4)
abline(lm(y4~x), col="green", lwd=2)

plot(lm(y4~x))

cor(x,y1)
cor(x, y2)

y<-rnorm(2001)
m<-lm(y~x)
summary(m)
res <- lm(y~x)$residual
cor(res,y)
cor(res, m$fitted)
```

What correlation does not measure: 
- whether variables are realted 
- whether a line is appropriate 
- "bang for buck"
- whether model is useful -- coefficients are always right for the mdoel that they are in, so their meaning and implications highly depend on the model. 


```{r}
fuel[1:4,]
plot(License, Fuel)
identify(License, Fuel, label=state)
pairs(cbind(Fuel, Tax, License))
cor(cbind(Fuel, Tax, License))

library(lattice)
levelplot(License~Tax*Fuel)

### need to install rgl in order to make this command run. 
##plot3d(Tax,License,Fuel)
```
Plotting in 3d not the most helpful. So what do we do?

```{r}
fuelstar<- Fuel/(License/100)
cor(Fuel, fuelstar)
summary(lm(fuelstar~Tax))
cor(fuelstar, Tax)
#cor(fuel,Tax)
#cor(Fuel,Tax)



```

$ Y= \beta_{0} +\beta_{1}Tax + \beta_{2}License +\epsilon$
$\epsilon ~ N(0, \sigma^2)$

pick $\hat\beta_{0}, \hat\beta_{1}, \hat\beta_{2}$ to minimize --> Now just solving three linear equations with three unknowns. 

General form of the linear model $ Y= \beta_{0} + \beta_{1}X_{1} + \beta_{2}X_{2} + ... + \beta_{k}X_{k} + \epsilon$ where $ \epsilon ~ N(0, \sigma^2)$


```{r}
lm(Fuel~Tax + License)

fuel[1:3,]
m0<-lm(Fuel~1)
m1<- lm(Fuel~Tax)
m2 <- lm(Fuel~Tax + License)
anova(m0, m1)
anova(m0, m2)
 
m2$resid
summary(m2)
```
**The coefficient of a variable depends upon what other variables are in the model** 


