---
title: "Notes, September 24th, More_Regression"
output: 
  html_document:
    keep_md: true
---

# Tuesday, September 24th 
## Mutliple Regression 
## Read Sheather chapters 6 (Regression by Stages) & 5 (General Linear Hypothesis)

### problem set out Thursday, 9/26. Due Tuesday Oct 15. 

```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
library(ppcor)
```


```{r}
vocabulary
attach(vocabulary)
summary(lm(Vocab~Age))
confint(lm(Vocab~Age))
plot(Age,Vocab, xlim=c(0, 7), ylim=c(-1000,3000))
abline(lm(Vocab~Age))
abline(h=0, col="brown")
agegt15<-pmax(0,Age-1.5)
agegt15
summary(lm(Vocab~Age+agegt15))
confint(lm(Vocab~Age+agegt15))

rbind(Age, agegt15)

plot(Age, Vocab)
points(Age, lm(Vocab~Age+agegt15)$fitted, pch=16, col="orange")

plot(lm(Vocab~Age))
detach(vocabulary)
```
agegt15 <- how much older than they are than 1.5 years. 

## Regression by Stages 

- Regress Fuel on Licenses, get residuals (fl)
- Regress Tax on on Licenses get residuals (tl)
- Regress residuals on residuals (fl on tl)
 - remove everything that Licenses can uniquely explain about Fuel and Tax. Then see what the residuals can explain. What's unique to Licenses that can't be predicted by Tax. 



```{r}
attach(fuel)
m2<-lm(Fuel~Tax+License)
m2

fl<- lm(Fuel~License)$residual
tl<- lm(Tax~License)$residual
lm(fl~tl)

ft<-lm(Fuel~Tax)$residual
lt<-lm(License~Tax)$residual
lm(ft~lt)

plot(tl, fl, xlab="Tax Residuals", ylab="Fuel Residuals")
abline(lm(fl~tl), col="blue")
avPlot(m2, Tax)

avPlot(m2,License)
avPlots(m2)
addedvarplot(lm(Fuel~License), Tax)

```

### Added Variable plot
abline(lm(fl~tl), col="blue") OR avPlot(m2, Tax)
 -> this is the line of the multiple regression (The Tax coefficient of m2)
 
### partial correlation: correlation between two residuals 

```{r}
cor(tl, fl)
cor(Tax, Fuel)

pcor(cbind(Fuel, Tax, License))


attach(partialcorEG)
pcor(partialcorEG)


m<-lm(Fuel~Tax+License+Inc)

iTL<-lm(Inc~Tax+License)$residual
fTL<-lm(Fuel~Tax+License)$residual
lm(fTL~iTL)
plot(iTL, fTL)
avPlot(lm(Fuel~Tax+License+Inc), Inc)
```

### General Linear Hypothesis  -- Chapter 5 in Sheather

$H_{0}:\beta =0$ look at t-test 
$H_{o}:\beta_{1}=\beta_{2}=0$ look at F-test 

C is a subset of of {1,2,3,...,k}, $H_{0}: \beta_{d}=0$ for j in C

Fit Full model all k variables (Tax, License, Inc)
Fit Reduced model 

```{r}
m0<-lm(Fuel~1)
m0
m1<-lm(Fuel~Tax)
anova(m0,m1)
anova(m0,m)
```





