---
title: "9_3_Probability"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
```

# Tuesday, September 3 
## Probability 

### confidence interval 
95% - 95% of the time we use it it will cover $/mu$ and 5% of the time $/mu$ is missed 
[A,B] for $/mu$

$\bar{x} \pm s$ where s is the standard deviation 
$\bar{x} \pm (t*s)/\sqrt(n)$


t- value from t table with n-1 degrees of freedom (d.f.) 


```{r}
bloodpressure$change
boxplot(bloodpressure$change)
sd(bloodpressure$change)
qt(.975, 10000)
t.test(17+rnorm(100))$conf.int
qnorm((1:15)/16)
#plot(abline(v=qnorm((1:15)/16)))
plot(qnorm((1:15)/16), sort(bloodpressure$change))
par(mfrow=c(2,2))
qqnorm(rnorm(15))
qqnorm(rnorm(100))
qqnorm(rnorm(500))
qqnorm(rnorm(5000))
par(mfrow=c(2,2))
qqnorm(rcauchy(100))
qqnorm(rcauchy(200))
qqnorm(rcauchy(100))
qqnorm(rcauchy(300))

```

### hypothesis testing 
#### an example 
Given 5 cards in a poker hand, whats the chance of a Royal Straight Flush

```{r}
possible_hands<-choose(52,5)
RSF <- 4/possible_hands
```
Chance that it will happen twice 
```{r}
RSF*RSF 
```
Chance that it will happen thrice 
```{r}
RSF*RSF*RSF
```

**The logic of hypothesis testing**: start with a hypothesis that we are not agnostic about. We have a preference. Compute a chance of it being true. 


Null Hypothesis: $H_{0}: \mu = \mu_{0}$ 
Alternative Hypothesis: $H_{A}: \mu \neq 0$

pt command: p-value, from t tables, chance that t is less than 1 standard deviation away from the mean. pt(s,d.f.)
multiply pt by two because we're doing a two-sided test. 

```{r}
pt(-8.12, 14)
pt(-8.12,14)*2
t.test(bloodpressure$change)

```


What if I tested every $\me_{0}$ 


