---
title: "9_26 Regression_cont"
output:
  html_document:
    keep_md: true
---

### FOR PROBLEMSET - REDOWNLOAD R WORKSPACE. NEED THE PAINTER DATA SET. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Thursday, September 26: General Linear Regression 
### Read Box's use and abuse of Regression


```{r}
load("/Users/isabellelangrock/Desktop/Stats500/Rst500.RData")
library(ltm)
library(car) 
library(ppcor)
```
## General Linear Hypothesis 
```{r}
attach(fuel)
mfull<-lm(Fuel~Tax+License+Inc)
mfull

```

Compare 2 nested models, one model eliminates variables in the other
full model all k vars
reduced model, rename varaibles (C) 

### Detailed Version
```{r}
mfull
m0<-lm(Fuel~1)
mreduced<- lm(Fuel~Tax)
anova(m0, mfull)
anova(m0, mreduced)
anova(mreduced)
1-pf(31.9,2,44)

```
## Short version 
```{r}
anova(mreduced,mfull)
```

We've been comparing to the null model (just the constant variable). 
We hypothesis that some of the coefficients are o, we fit the full model that assumes that's not true, fit the reduced model, look at the sum of squares for error --> has it reduced enough to justify using the variables (it'll always be smaller for the full model but we want to make sure it's small enough)


Now, testing the hypothesis $H_{0}: \beta_{tax}=0$
```{r}
mreduced2<-lm(Fuel~License+Inc)
anova(m0, mfull)
anova(m0,mreduced2)
anova(mreduced2, mfull)
summary(mfull)
```

Is there a sum of squares for a group of variables? no! 

##Review of Logorithms 

$10^3$= 10 x 10 x 10 = 1000 
$log_{10}(1000)= 3$
$y=10^a  <-> log_{10}(y)=a$
$y=b^a <-> log_{b}(y)=a$
$log_{2}(y) = \frac {log_{10}(y)}{log_{10}(2)}$
```{r}
log10(1000)
10^3
log2(8)
2^3
log10(8)/log10(2)
log(8)
```

What base should we use? e= 2.71828...

$y=e^a <=> log_{e}(y)= ln(y) = log(y)= a$
$\frac {d e^x}{d x} = e^x$
$ \frac {d ln(y)}{d y} = \frac {1}{y}$
ln(y) = natural log = $log_{e}$

Never write $e^{ugly}$ always write exp(ugly). 

Properties of log: 
- log(x) defined for x>0
- log(xy) = log(x) + log(y)
- $log(x^a)$= a log(x)
- x<y <=> log(x) < log(y)
- exp(x+y) = exp(x)exp(y)
- exp(log(y))=log(exp(y))=y
- 
- 



